---
title: "DS_Project"
output: pdf_document
date: "2023-03-20"
---

```{r, include = TRUE, warning = FALSE, message = FALSE}
library(tidyverse)
library(modelr)
library(lubridate)
library(dplyr)

economy_data = read_csv("Flights Price Prediction Dataset\\economy.csv", show_col_types = FALSE)
business_data = read_csv("Flights Price Prediction Dataset\\business.csv", show_col_types = FALSE)
flight = read_csv("Flights Price Prediction Dataset\\Clean_Dataset.csv", show_col_types = FALSE)
economy_data$time_taken = lubridate::hm(economy_data$time_taken)
business_data$time_taken = lubridate::hm(business_data$time_taken)
head(economy_data)
head(business_data)

flight["stops"][flight["stops"]== "zero"] <- '0'
flight["stops"][flight["stops"]== "one"] <- '1'
flight["stops"][flight["stops"]== "two_or_more"] <- '2'
flight$stops = as.numeric(flight$stops)
typeof(flight$stops)
head(flight)
flight$sector = paste(flight$source_city, flight$destination_city)
flight$date <- as.Date(flight$date)
flight$weekday <- strftime(flight$date,"%A") 
head(flight)

```

##distribution with respect to departure time


```{r, warning = FALSE, message = FALSE, fig.width = 10, fig.height = 3}
#distribution with respect to departure time

ggplot (data = economy_data) +
  geom_histogram(mapping = aes(x=dep_time))

ggplot (data = business_data) +
  geom_histogram(mapping = aes(x=dep_time))
```

##distribution with respect to duration


```{r, warning = FALSE, message = FALSE, fig.width = 10, fig.height = 3}
ggplot (data = economy_data) +
  geom_histogram(mapping = aes(x=(lubridate::as.duration(time_taken)/3600)))

ggplot (data = business_data) +
  geom_histogram(mapping = aes(x=(lubridate::as.duration(time_taken)/3600)))
```



##Price distribution based on the day of the week


```{r, warning = FALSE, message = FALSE, fig.width = 10, fig.height = 3}
  #Converting date to weekday
economy_data$date <- as.Date(economy_data$date)
economy_data$weekday <- strftime(economy_data$date,"%A") 
business_data$date <- as.Date(business_data$date)
business_data$weekday <- strftime(business_data$date,"%A")
  #Plotting price vs day of economy
weekday_graph_economy <- ggplot (data = economy_data) +
  geom_point(mapping = aes (x = weekday, y = price))+
  geom_point(mapping = aes (x = weekday, y = median(price)),color="red")+
  geom_point(mapping = aes (x = weekday, y = mean(price)),color="gold2")
(weekday_graph_economy)
  #Plotting price vs day of business
weekday_graph_business <- ggplot (data = business_data) +
  geom_point(mapping = aes (x = weekday, y = price))+
  geom_point(mapping = aes (x = weekday, y = median(price)),color="red")+
  geom_point(mapping = aes (x = weekday, y = mean(price)),color="gold2")
(weekday_graph_business)
  #Grouping dataset by weekday
eco_grp_weekday = economy_data %>% group_by(weekday)  %>%
  summarise(mean_price = mean(price),
            median_price = median(price),
            .groups = 'drop')
bus_grp_weekday = business_data %>% group_by(weekday)  %>%
  summarise(mean_price = mean(price),
            median_price = median(price),
            .groups = 'drop')
    #Mean and median price in economy by weekday
grp_economy <- ggplot (data = eco_grp_weekday) +
  geom_point(mapping = aes (x = weekday, y = mean_price),color="blue")+
  geom_point(mapping = aes (x = weekday, y = median_price),color="red")+
  ylab("price")
(grp_economy)
    #Mean and median price in business by weekday
grp_business <- ggplot (data = bus_grp_weekday) +
  geom_point(mapping = aes (x = weekday, y = mean_price),color="blue")+
  geom_point(mapping = aes (x = weekday, y = median_price),color="red")+
  ylab("price")
(grp_business)
```


# Price distribution on the basis of time in the day (currently merged business and economy)

```{r, warning = FALSE, message = FALSE, fig.width = 10, fig.height = 3}
ggplot(data = flight)+
  geom_point(mapping = aes(departure_time, price))+
  labs(y= "Price in INR", x = "Time of Departure")
```


# Price distrbution on the basis of duration of the flight (Economy & Business)


```{r, warning = FALSE, message = FALSE, fig.width = 10, fig.height = 3}
ggplot (data = economy_data) +
  geom_point(mapping = aes(x=(lubridate::as.duration(time_taken)/3600), y=price)) +
  labs(y= "Price in INR", x="Flight duration", title = "Economy class") +
  geom_smooth(mapping = aes(x=(lubridate::as.duration(time_taken)/3600), y=price))

ggplot (data = business_data) +
  geom_point(mapping = aes(x=(lubridate::as.duration(time_taken)/3600), y=price)) +
  labs(y= "Price in INR", x="Flight duration", title = "Business class") +
  geom_smooth(mapping = aes(x=(lubridate::as.duration(time_taken)/3600), y=price))



```
# Price distrbution on the basis of duration of the flight (Economy & Business)


```{r, warning = FALSE, message = FALSE, fig.width = 10, fig.height = 3}
library(tidyverse)
library(glmnet)
library(caret)
ggplot (data = economy_data) +
  geom_point(mapping = aes(x=(lubridate::as.duration(time_taken)/3600), y=price)) +
  labs(y= "Price in INR", x="Flight duration", title = "Economy class") +
  geom_smooth(mapping = aes(x=(lubridate::as.duration(time_taken)/3600), y=price))

ggplot (data = business_data) +
  geom_point(mapping = aes(x=(lubridate::as.duration(time_taken)/3600), y=price)) +
  labs(y= "Price in INR", x="Flight duration", title = "Business class") +
  geom_smooth(mapping = aes(x=(lubridate::as.duration(time_taken)/3600), y=price))
```


```{r, warning = FALSE, message = FALSE, fig.width = 10, fig.height = 3}
y = flight$price
x = model.matrix( ~ ., data =(select(flight,-price, -flight, -source_city, -destination_city, -date)))

index = sample(nrow(x), floor(0.9 * nrow(x)))
x_train = x[index, ]
y_train = y[index]
x_test = x[-index, ]
y_test = y[-index]

model = cv.glmnet(x_train,y_train,alpha =1,folds = 5)
optimal_lambda = model$lambda.min
optimal_model = glmnet(x_train,y_train,alpha =1,lambda = optimal_lambda)
summary(optimal_model)
coef(optimal_model)
```

```{r, warning = FALSE, message = FALSE, fig.width = 10, fig.height = 3}
(c = coef(optimal_model))
(sum(c==0))
unique(flight$airline)
```

```{r, warning = FALSE, message = FALSE, fig.width = 10, fig.height = 3}
library(pROC)

predicted = predict(optimal_model, x_test)
auc(y_test, predicted)
postResample(y_test, predicted)


```
