---
title: "DS_Project"
output: pdf_document
date: "2023-03-20"
---

```{r, include = TRUE, warning = FALSE, message = FALSE}
library(tidyverse)
library(modelr)
library(lubridate)
library(dplyr)
library(glmnet)
library(caret)
library(pROC)
library(caTools)
library(randomForest)
#library(AICcmodavg)

economy_data = read_csv("Flights Price Prediction Dataset\\economy.csv", show_col_types = FALSE)
business_data = read_csv("Flights Price Prediction Dataset\\business.csv", show_col_types = FALSE)
flight = read_csv("Flights Price Prediction Dataset\\Clean_Dataset.csv", show_col_types = FALSE)
flight_new = read_csv("Flights Price Prediction Dataset\\Cleaned_dataset.csv", show_col_types = FALSE)
economy_data$time_taken = lubridate::hm(economy_data$time_taken)
business_data$time_taken = lubridate::hm(business_data$time_taken)
head(economy_data)
head(business_data)


flight["stops"][flight["stops"]== "zero"] <- '0'
flight["stops"][flight["stops"]== "one"] <- '1'
flight["stops"][flight["stops"]== "two_or_more"] <- '2'
flight$stops = as.numeric(flight$stops)
typeof(flight$stops)
head(flight)
flight$sector = paste(flight$source_city, flight$destination_city)
flight$date <- as.Date(flight$date)
flight$weekday <- strftime(flight$date,"%A") 
head(flight)
head(flight_new)
unique(flight_new$Total_stops)
flight_new["Total_stops"][flight_new["Total_stops"]== "non-stop"] <- '0'
flight_new["Total_stops"][flight_new["Total_stops"]== "1-stop"] <- '1'
flight_new["Total_stops"][flight_new["Total_stops"]== "2+-stop"] <- '2'
flight_new$Total_stops = as.numeric(flight_new$Total_stops)
flight_new$sector = paste(flight_new$Source, flight_new$Destination)
head(flight_new)
```

##Distribution of data with respect to departure time


```{r, warning = FALSE, message = FALSE, fig.width = 10, fig.height = 3}
#distribution with respect to departure time

ggplot (data = economy_data) +
  geom_histogram(mapping = aes(x=dep_time))

ggplot (data = business_data) +
  geom_histogram(mapping = aes(x=dep_time))
```

##Distribution of data with respect to duration


```{r, warning = FALSE, message = FALSE, fig.width = 10, fig.height = 3}
ggplot (data = economy_data) +
  geom_histogram(mapping = aes(x=(lubridate::as.duration(time_taken)/3600)))

ggplot (data = business_data) +
  geom_histogram(mapping = aes(x=(lubridate::as.duration(time_taken)/3600)))
```



##Price distribution based on the day of the week


```{r, warning = FALSE, message = FALSE, fig.width = 10, fig.height = 3}
  #Converting date to weekday
economy_data$date <- as.Date(economy_data$date)
economy_data$weekday <- strftime(economy_data$date,"%A") 
business_data$date <- as.Date(business_data$date)
business_data$weekday <- strftime(business_data$date,"%A")
  #Plotting price vs day of economy
weekday_graph_economy <- ggplot (data = economy_data) +
  geom_point(mapping = aes (x = weekday, y = price))+
  geom_point(mapping = aes (x = weekday, y = median(price)),color="red")+
  geom_point(mapping = aes (x = weekday, y = mean(price)),color="gold2")
(weekday_graph_economy)
  #Plotting price vs day of business
weekday_graph_business <- ggplot (data = business_data) +
  geom_point(mapping = aes (x = weekday, y = price))+
  geom_point(mapping = aes (x = weekday, y = median(price)),color="red")+
  geom_point(mapping = aes (x = weekday, y = mean(price)),color="gold2")
(weekday_graph_business)
  #Grouping dataset by weekday
eco_grp_weekday = economy_data %>% group_by(weekday)  %>%
  summarise(mean_price = mean(price),
            median_price = median(price),
            .groups = 'drop')
bus_grp_weekday = business_data %>% group_by(weekday)  %>%
  summarise(mean_price = mean(price),
            median_price = median(price),
            .groups = 'drop')
    #Mean and median price in economy by weekday
grp_economy <- ggplot (data = eco_grp_weekday) +
  geom_point(mapping = aes (x = weekday, y = mean_price),color="blue")+
  geom_point(mapping = aes (x = weekday, y = median_price),color="red")+
  ylab("price")
(grp_economy)
    #Mean and median price in business by weekday
grp_business <- ggplot (data = bus_grp_weekday) +
  geom_point(mapping = aes (x = weekday, y = mean_price),color="blue")+
  geom_point(mapping = aes (x = weekday, y = median_price),color="red")+
  ylab("price")
(grp_business)
```


##Price distribution on the basis of time in the day (currently merged business and economy)

```{r, warning = FALSE, message = FALSE, fig.width = 10, fig.height = 3}
ggplot(data = flight)+
  geom_point(mapping = aes(departure_time, price))+
  labs(y= "Price in INR", x = "Time of Departure")
```


##Price distrbution on the basis of duration of the flight (Economy & Business)


```{r, warning = FALSE, message = FALSE, fig.width = 10, fig.height = 3}
ggplot (data = economy_data) +
  geom_point(mapping = aes(x=(lubridate::as.duration(time_taken)/3600), y=price)) +
  labs(y= "Price in INR", x="Flight duration", title = "Economy class") +
  geom_smooth(mapping = aes(x=(lubridate::as.duration(time_taken)/3600), y=price))

ggplot (data = business_data) +
  geom_point(mapping = aes(x=(lubridate::as.duration(time_taken)/3600), y=price)) +
  labs(y= "Price in INR", x="Flight duration", title = "Business class") +
  geom_smooth(mapping = aes(x=(lubridate::as.duration(time_taken)/3600), y=price))



```
##Price distrbution on the basis of duration of the flight (Economy & Business)


```{r, warning = FALSE, message = FALSE, fig.width = 10, fig.height = 3}

ggplot (data = economy_data) +
  geom_point(mapping = aes(x=(lubridate::as.duration(time_taken)/3600), y=price)) +
  labs(y= "Price in INR", x="Flight duration", title = "Economy class") +
  geom_smooth(mapping = aes(x=(lubridate::as.duration(time_taken)/3600), y=price))

ggplot (data = business_data) +
  geom_point(mapping = aes(x=(lubridate::as.duration(time_taken)/3600), y=price)) +
  labs(y= "Price in INR", x="Flight duration", title = "Business class") +
  geom_smooth(mapping = aes(x=(lubridate::as.duration(time_taken)/3600), y=price))
```

##Model Fitting: Lasso Model
```{r, warning = FALSE, message = FALSE, fig.width = 10, fig.height = 3}
#Defining predictor and response variables
y = flight$price
x = model.matrix( ~ ., data =(select(flight,-price, -flight, -source_city, -destination_city, -date)))

#Splitting data
index = sample(nrow(x), floor(0.8 * nrow(x)))
x_train = x[index, ]
y_train = y[index]
x_test = x[-index, ]
y_test = y[-index]

#Fitting the model
model = cv.glmnet(x_train,y_train,alpha =1,folds = 5)
optimal_lambda = model$lambda.min
optimal_model = glmnet(x_train,y_train,alpha =1,lambda = optimal_lambda)
summary(optimal_model)

#Extracting model coefficients
coef(optimal_model)
```

```{r, warning = FALSE, message = FALSE, fig.width = 10, fig.height = 3}
#Counting number of variables removed in the model
(c = coef(optimal_model))
(sum(c==0))

```

```{r, warning = FALSE, message = FALSE, fig.width = 10, fig.height = 3}
#Extracting pmodel predictions
predicted = predict(optimal_model, x_test)

#Evaluation metrics
auc(y_test, predicted)
postResample(y_test, predicted)

```

##Model Fitting: Lasso Model to new data
```{r, warning = FALSE, message = FALSE, fig.width = 10, fig.height = 3}
#Defining predictor and response variables
head(flight_new)
y = flight_new$Fare
x = model.matrix( ~ ., data =(select(flight_new,-Fare, -Flight_code, -Source, -Destination, -Date_of_journey)))

#Splitting data
index = sample(nrow(x), floor(0.8 * nrow(x)))
x_train = x[index, ]
y_train = y[index]
x_test = x[-index, ]
y_test = y[-index]

#Fitting the model
model = cv.glmnet(x_train,y_train,alpha =1,folds = 5)
optimal_lambda = model$lambda.min
optimal_model = glmnet(x_train,y_train,alpha =1,lambda = optimal_lambda)
summary(optimal_model)

#Extracting model coefficients
coef(optimal_model)
```
```{r, warning = FALSE, message = FALSE, fig.width = 10, fig.height = 3}
#Extracting pmodel predictions
predicted = predict(optimal_model, x_test)

#Evaluation metrics
auc(y_test, predicted)
postResample(y_test, predicted)

```
##Model Fitting: Lasso Model to new data
```{r, warning = FALSE, message = FALSE, fig.width = 10, fig.height = 3}
#Defining predictor and response variables
head(flight_new)
y = flight_new$Fare
x = model.matrix( ~ ., data =(select(flight_new,-Fare, -Flight_code, -sector, -Date_of_journey, -Duration_in_hours  )))

#Splitting data
index = sample(nrow(x), floor(0.8 * nrow(x)))
x_train = x[index, ]
y_train = y[index]
x_test = x[-index, ]
y_test = y[-index]

#Fitting the model
model = cv.glmnet(x_train,y_train,alpha =1,folds = 5)
optimal_lambda = model$lambda.min
optimal_model = glmnet(x_train,y_train,alpha =1,lambda = optimal_lambda)
summary(optimal_model)

#Extracting model coefficients
coef(optimal_model)
```

```{r, warning = FALSE, message = FALSE, fig.width = 10, fig.height = 3}
#Extracting pmodel predictions
predicted = predict(optimal_model, x_test)
residuals = predicted - y_test
plot(y_test,residuals)

#Evaluation metrics
auc(y_test, predicted)
postResample(y_test, predicted)

```

```{r, warning = FALSE, message = FALSE, fig.width = 10, fig.height = 3}

#Defining predictor and response variables
head(flight_new)
y = flight_new$Fare
x = model.matrix( ~ ., data =(select(flight_new,-Fare, -Flight_code, -sector, -Date_of_journey, -Duration_in_hours)))
head(x)
#Splitting data
index = sample(nrow(x), floor(0.7 * nrow(x)))
x_train = x[index, ]
y_train = y[index]
x_test = x[-index, ]
y_test = y[-index]

classifier_RF = randomForest(x = x_train, y = y_train, ntree = 10, do.trace=TRUE)
classifier_RF

predicted = predict(classifier_RF, x_test)
residuals = predicted - y_test
plot(y_test,residuals)
#Evaluation metrics
postResample(y_test, predicted)
saveRDS(classifier_RF, file = "model_file_Random_Forest.rds")

model <- readRDS("model_file_Random_Forest.rds")
predicted = predict(model, x_test)
residuals = predicted - y_test
plot(y_test,residuals)
#Evaluation metrics
postResample(y_test, predicted)
head(x_test)
```
```{r, warning = FALSE, message = FALSE, fig.width = 10, fig.height = 3}

#Defining predictor and response variables
head(flight_new)
y = flight_new$Fare
x = model.matrix( ~ ., data =(select(flight_new,-Fare, -Flight_code, -sector, -Date_of_journey, -Duration_in_hours)))
head(x)
#Splitting data
index = sample(nrow(x), floor(0.7 * nrow(x)))
x_train = x[index, ]
y_train = y[index]
x_test = x[-index, ]
y_test = y[-index]

classifier_RF = randomForest(x = x_train, y = y_train, ntree = 25, do.trace=TRUE)
classifier_RF

predicted = predict(classifier_RF, x_test)
residuals = predicted - y_test
plot(y_test,residuals)
#Evaluation metrics
postResample(y_test, predicted)
saveRDS(classifier_RF, file = "model_file_Random_Forest_25_1.rds")

model <- readRDS("model_file_Random_Forest_25_1.rds")
predicted = predict(model, x_test)
residuals = predicted - y_test
plot(y_test,residuals)
#Evaluation metrics
postResample(y_test, predicted)
head(x_test)

```

```{r, warning = FALSE, message = FALSE, fig.width = 10, fig.height = 4}
library(randomForestExplainer)

model <- readRDS("model_file_Random_Forest_25_1.rds")
predicted = predict(model, x_test)
residuals = predicted - y_test
plot(y_test,residuals)
#Evaluation metrics
model
postResample(y_test, predicted)

plot(model)

varImpPlot(model) 


```

```{r, warning = FALSE, message = FALSE, fig.width = 10, fig.height = 3}
unique(flight_new$Airline)
#head(flight_new)
unique(flight_new$Class)
unique(flight_new$Arrival)
unique(flight_new$Destination)

```
```{r}

Airline = "SpiceJet"
Source = "Delhi"
Destination = "Mumbai"
Flight_date = as.Date("11/02/22") 
Days_left = 5
Class = "Economy" 
Total_stops = 1
Arrival = "Before 6 AM" 
Departure = "6 AM - 12 PM"


Journey_day = strftime(Flight_date,"%A") 
flight_inpt = data.frame(Journey_day, Airline, Class, Source, Departure, Total_stops, Arrival, Destination, Days_left)
inp = model.matrix( ~ ., data =rbind(flight_inpt, select(flight_new,-Fare, -Flight_code, -sector, -Date_of_journey, -Duration_in_hours)))
prd = predict(optimal_model, inp[1, ])
prd
x_test[1, ]
inp[1, ]
flight_inpt
select(flight_new,-Fare, -Flight_code, -sector, -Date_of_journey, -Duration_in_hours)
class(prd)
```


```{r, warning = FALSE, message = FALSE, fig.width = 10, fig.height = 3}

flightPricePredict <- function(Airline, Source, Destination, Flight_date, Days_left, Class, Total_stops, Arrival, Departure, Duration_in_hours = 2){
  Journey_day = strftime(Flight_date,"%A") 
  flight_inpt = data.frame(Journey_day, Airline, Class, Source, Departure, Total_stops, Arrival, Destination, Duration_in_hours, Days_left)
  inp = model.matrix( ~ ., data =rbind(flight_inpt, select(flight_new,-Fare, -Flight_code, -sector, -Date_of_journey)))
  prd = predict(optimal_model, inp[1, ])
  return(prd)
}
flightPricePredict("SpiceJet", "Delhi", "Mumbai", as.Date("11/02/22"), 5, "Economy", 0, "Before 6 AM", "6 AM - 12 PM") 

```

```{r, warning = FALSE, message = FALSE, fig.width = 10, fig.height = 5}
flight_new %>% 
  ggplot () +
  geom_bar(mapping = aes(x=Airline))

flight_new %>% 
  ggplot () +
  geom_bar(mapping = aes(x=Class))

flight_new %>% 
  ggplot () +
  geom_bar(mapping = aes(x=Journey_day))

ggplot (data = flight_new) +
  geom_bar(mapping = aes(x=Journey_day, y=Fare),stat = "summary", fun = "mean")

ggplot (data = flight_new) +
  geom_bar(mapping = aes(x=Days_left, y=Fare),stat = "summary", fun = "mean")

ggplot (data = flight_new) +
  geom_bar(mapping = aes(x=Airline, y=Fare),stat = "summary", fun = "mean")

ggplot (data = flight_new) +
  geom_bar(mapping = aes(x=Departure, y=Fare),stat = "summary", fun = "mean")

ggplot (data = flight_new) +
  geom_bar(mapping = aes(x=Arrival, y=Fare),stat = "summary", fun = "mean")

ggplot (data = flight_new) +
  geom_bar(mapping = aes(x=Source, y=Fare),stat = "summary", fun = "mean")

ggplot (data = flight_new) +
  geom_bar(mapping = aes(x=Destination, y=Fare),stat = "summary", fun = "mean")

ggplot (data = flight_new) +
  geom_bar(mapping = aes(x=Class, y=Fare),stat = "summary", fun = "mean")

ggplot (data = flight_new) +
  geom_bar(mapping = aes(x=Total_stops, y=Fare),stat = "summary", fun = "mean")

ggplot (data = flight_new) +
  geom_bar(mapping = aes(x=sector, y=Fare),stat = "summary", fun = "mean")+
  theme(axis.text.x=element_text(angle=90,hjust=1,vjust=0.5))


ggplot (data = flight_new) +
  geom_point(mapping = aes(x=Duration_in_hours, y=Fare)) +
  labs(y= "Price in INR", x="Flight duration") +
  geom_smooth(mapping = aes(x=Duration_in_hours, y=Fare))



```